---
title: "Rhesus scRNA Seurat Clustering"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(Seurat)
library(patchwork)
```

# Load the dataset
```{r}
# Load the dataset
scRNA_3_1.data <- Read10X(data.dir = "cellranger/3_1/outs/filtered_feature_bc_matrix")

# Initialize the Seurat object with the raw (non-normalized data).
scRNA_3_1 <- CreateSeuratObject(counts = scRNA_3_1.data, project = "rhesus", min.cells = 10, min.features = 200)
```

# Standard pre-processing workflow
```{r}
# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
scRNA_3_1[["percent.mt"]] <- PercentageFeatureSet(scRNA_3_1, pattern = "^MT-")

# Visualize QC metrics as a violin plot
VlnPlot(scRNA_3_1, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

# FeatureScatter is typically used to visualize feature-feature relationships, but can be used for anything calculated by the object, i.e. columns in object metadata, PC scores etc.
plot1 <- FeatureScatter(scRNA_3_1, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(scRNA_3_1, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2

scRNA_3_1_subset <- subset(scRNA_3_1, subset = nFeature_RNA > 200 & nFeature_RNA < 2500 & percent.mt < 5)
```

# Normalizing the data
```{r}
scRNA_3_1_normalized <- NormalizeData(scRNA_3_1_subset, normalization.method = "LogNormalize", scale.factor = 10000)
scRNA_3_1_features <- FindVariableFeatures(scRNA_3_1_normalized, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(scRNA_3_1_features), 10)

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(scRNA_3_1_features)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1
plot2

# Scaling the data
all.genes <- rownames(scRNA_3_1_features)
scRNA_3_1_scaled <- ScaleData(scRNA_3_1_features, features = all.genes)

# Perform linear dimensional reduction
scRNA_3_1_linear_reduction <- RunPCA(scRNA_3_1_scaled, features = VariableFeatures(object = scRNA_3_1_scaled))
```

# Examine and visualize PCA results a few different ways
```{r}
print(scRNA_3_1_linear_reduction[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(scRNA_3_1_linear_reduction, dims = 1:2, reduction = "pca")

DimPlot(scRNA_3_1_linear_reduction, reduction = "pca")

DimHeatmap(scRNA_3_1_linear_reduction, dims = 1, cells = 500, balanced = TRUE)

DimHeatmap(scRNA_3_1_linear_reduction, dims = 1:15, cells = 500, balanced = TRUE)
```

# Determine the 'dimensionality' of the dataset
```{r}
# NOTE: This process can take a long time for big datasets, comment out for expediency. More approximate techniques such as those implemented in ElbowPlot() can be used to reduce computation time
scRNA_3_1_jackstraw <- JackStraw(scRNA_3_1_linear_reduction, num.replicate = 100)
scRNA_3_1_jackstraw <- ScoreJackStraw(scRNA_3_1_jackstraw, dims = 1:20)

JackStrawPlot(scRNA_3_1_jackstraw, dims = 1:15)

ElbowPlot(scRNA_3_1_jackstraw)
```

# Cluster the cells
```{r}
scRNA_3_1_cluster <- FindNeighbors(scRNA_3_1_jackstraw, dims = 1:10)
scRNA_3_1_cluster <- FindClusters(scRNA_3_1_cluster, resolution = 0.5)

# Look at cluster IDs of the first 5 cells
head(Idents(scRNA_3_1_cluster), 5)

# Run non-linear dimensional reduction (UMAP/tSNE)
scRNA_3_1_cluster <- RunUMAP(scRNA_3_1_cluster, dims = 1:10)
# note that you can set `label = TRUE` or use the LabelClusters function to help label individual clusters
DimPlot(scRNA_3_1_cluster, reduction = "umap")

# You can save the object at this point so that it can easily be loaded back in without having to rerun the computationally intensive steps performed above, or easily shared with collaborators.
saveRDS(scRNA_3_1_cluster, file = "scRNA_3_1_cluster_load.rds")
```

# Finding differentially expressed features (cluster biomarkers)
```{r}
# find all markers of cluster 1
cluster1.markers <- FindMarkers(scRNA_3_1_cluster, ident.1 = 1, min.pct = 0.25)
head(cluster1.markers, n = 5)

# find markers for every cluster compared to all remaining cells, report only the positive ones
markers_3_1 <- FindAllMarkers(scRNA_3_1_cluster, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
markers_3_1 %>% group_by(cluster) %>% top_n(n = 2, wt = avg_logFC)

# Use cell marker databases listed in github repository to determine cell type: https://github.com/aedlim/rhesus_scrna

# Use FeaturePlot to visualize certain gene markers
#Cluster0
FeaturePlot(scRNA_3_1_cluster, features = c("DLX6-AS1", "ERBB4"))
#Cluster1
FeaturePlot(scRNA_3_1_cluster, features = c("PPP1R17", "HBA1"))
#Cluster2
FeaturePlot(scRNA_3_1_cluster, features = c("SATB2", "CSRP2"))
#Cluster3
FeaturePlot(scRNA_3_1_cluster, features = c("NEUROD2", "SLA"))
#Cluster4
FeaturePlot(scRNA_3_1_cluster, features = c("MEF2C", "RPRM"))
#Cluster5
FeaturePlot(scRNA_3_1_cluster, features = c("UBE2C", "KIAA0101"))
#Cluster6
FeaturePlot(scRNA_3_1_cluster, features = c("ZFHX3", "FOXP1"))
#Cluster7
FeaturePlot(scRNA_3_1_cluster, features = c("NRN1", "NFIB"))
#Cluster8
FeaturePlot(scRNA_3_1_cluster, features = c("PTN", "VIM"))
#Cluster9
FeaturePlot(scRNA_3_1_cluster, features = c("ANK3", "RALYL"))
#Cluster10
FeaturePlot(scRNA_3_1_cluster, features = c("SPP1", "LGALS3"))
#Cluster11
FeaturePlot(scRNA_3_1_cluster, features = c("OLIG1", "SCRG1"))
#Cluster12
FeaturePlot(scRNA_3_1_cluster, features = c("RAMP2", "RNASE1"))
```
#Assigning cell type identity to clusters
```{r}
new.cluster.ids_3_1 <- c("Interneurons", "Intermediate progenitor cells", "Maturing excitatory neuron", "Newborn excitatory neuron", "Dopaminergic amacrine neuron", "Neural stem cell", "Striatal inhibitory neuron", "Cortical neuron", "Radial glia", "Pyramidal neuron","Microglia", "Oligodendrocyte progenitor cells", "Pericyte")
names(new.cluster.ids_3_1) <- levels(scRNA_3_1_cluster)
scRNA_3_1_cluster <- RenameIdents(scRNA_3_1_cluster, new.cluster.ids_3_1)
DimPlot(scRNA_3_1_cluster, reduction = "umap", label = FALSE, pt.size = 0.5)
DimPlot(scRNA_3_1_cluster, reduction = "umap", label = TRUE, pt.size = 0.5)+NoLegend()
```

#Output number of cells in each cluster
```{r}
table(scRNA_3_1_cluster@active.ident)
```

# Save final object for easy loading
```{r}
saveRDS(scRNA_3_1_cluster, file = "scRNA_3_1_cluster_final.rds")
```
